---
title: "Analyzing E-commerce Consumer Behaviour"
author: "Phil (Fei Han) & Doniyorbek Komi"
format: pdf
editor: visual
date: "2026-03-19"

---
```{r}
# Load libraries
library(tidyverse)
library(corrplot)
library(viridis)
library(scales)
library(car)
library(pROC)

# --- Clean and Preprocess Data ---
df_clean <- read.csv("online_shoppers_intention.csv") %>%
  mutate(
    # Convert Revenue to Factor
    Revenue = factor(Revenue, levels = c(FALSE, TRUE), labels = c("No Purchase", "Purchase")),
    
    # Convert Month (Define Order: Feb is the Reference Level)
    Month = factor(Month, levels = c("Feb", "Mar", "May", "June", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")),
    
    # Convert VisitorType and Weekend
    VisitorType = as.factor(VisitorType),
    Weekend = factor(Weekend, levels = c(FALSE, TRUE), labels = c("Weekday", "Weekend")),
    
    # Create Numeric Label for calculations
    is_purchase = as.numeric(Revenue == "Purchase")
  ) %>%
  drop_na()

# Define a consistent theme for the project
project_theme <- theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(size = 12),
    legend.position = "bottom"
  )
```

Exploratory Data Analysis (EDA)
distribution of Page Values
We begin by examining our primary predictors. Our initial hypothesis suggests that PageValues significantly influences the likelihood of a transaction.
```{r}
#| label: data-summary
#| message: false

# Check for missing values
cat("Missing values in dataset:", sum(is.na(df_clean)), "\n")

# View the distribution of our target variable: Revenue
table(df_clean$Revenue)
```
```{r}
#| label: fig-pagevalues
#| fig-cap: "Comparison of Page Values for Purchase vs. No Purchase"

ggplot(df_clean, aes(x = Revenue, y = PageValues, fill = Revenue)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#00CC66")) +
  labs(
    title = "Does Page Value Predict Purchase?",
    subtitle = "Higher PageValues are strongly associated with successful transactions",
    x = "Transaction Outcome (Revenue)",
    y = "Page Value Score"
  ) +
  project_theme
```

```{r}
#| label: fig-monthly-trend
#| fig-cap: "Monthly Conversion Rates Throughout the Year"

df_clean %>% 
  group_by(Month) %>% 
  summarise(Conversion_Rate = mean(Revenue == "Purchase")) %>% 
  ggplot(aes(x = Month, y = Conversion_Rate, group = 1)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(size = 3, color = "darkblue") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "E-commerce Seasonality: Monthly Conversion Rates",
    x = "Month of the Year",
    y = "Conversion Rate (%)"
  ) +
  project_theme
```
```{r}
#| label: fig-visitor-type
#| fig-cap: "Purchase Proportion by Visitor Type"

ggplot(df_clean, aes(x = VisitorType, fill = Revenue)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Visitor Behaviour: New vs. Returning",
    x = "Visitor Category",
    y = "Percentage",
    fill = "Purchased?"
  ) +
  project_theme
```
```{r}
#| label: fig-correlation
#| fig-height: 7
#| fig-width: 7

# Select numeric columns and calculate correlation
numeric_df <- df_clean %>% select(where(is.numeric))
cor_matrix <- cor(numeric_df)

corrplot(cor_matrix, method = "circle", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         title = "Correlation Heatmap of Website Metrics",
         mar = c(0,0,1,0))
```

```{r}
#| label: data-cleaning
#| message: false

df <- read.csv("online_shoppers_intention.csv")

df_clean <- df %>%
  mutate(
    Revenue = factor(Revenue, 
                     levels = c(FALSE, TRUE), 
                     labels = c("No Purchase", "Purchase")),
    
    Month = factor(Month, 
                   levels = c("Feb", "Mar", "May", "June", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")),
    
    VisitorType = as.factor(VisitorType),
    Weekend = factor(Weekend, 
                     levels = c(FALSE, TRUE), 
                     labels = c("Weekday", "Weekend")),
    
    is_purchase = as.numeric(Revenue == "Purchase")
  ) %>%
  drop_na()

summary(df_clean)
```


```{r}
# Create binary numeric response variable
df_clean$Revenue_num <- as.numeric(df_clean$Revenue == "Purchase")

# --- Step 1: Variable Selection (unweighted) ---
predictors <- setdiff(names(df_clean), c("Revenue", "Revenue_num", "case_weight", "is_purchase"))
clean_formula <- as.formula(paste("Revenue_num ~", paste(predictors, collapse = " + ")))

step_model_base <- glm(clean_formula, data = df_clean, family = binomial)
aic_var_selection <- step(step_model_base, direction = "both", trace = 0)

# --- Step 2: Final Model ---
final_formula <- formula(aic_var_selection)
aic_model <- glm(final_formula, data = df_clean, family = binomial)
summary(aic_model)
```
```{r}
# --- Step 1: Manually define predictors (exclude auxiliary columns) ---
# Ensure the model only sees raw features, not weight or helper columns
predictors <- setdiff(names(df_clean), c("Revenue", "case_weight", "is_purchase"))
clean_formula <- as.formula(paste("Revenue ~", paste(predictors, collapse = " + ")))

# --- Step 2: Fit base model (for variable selection only) ---
# No weights at this stage — keeps AIC logic clean and avoids errors in step()
step_model_base <- glm(clean_formula, data = df_clean, family = binomial)

# --- Step 3: Stepwise variable selection ---
# trace = 0 suppresses the selection log, keeping the rendered document clean
aic_var_selection <- step(step_model_base, direction = "both", trace = 0)

# --- Step 4: Fit final weighted model ---
# Extract the selected formula from aic_var_selection
final_formula <- formula(aic_var_selection)

# Key fix: use df_clean$case_weight to explicitly reference the column
aic_model <- glm(final_formula, 
                 data    = df_clean, 
                 family  = quasibinomial, 
                 weights = df_clean$case_weight)

# View results
summary(aic_model)
```


```{r}
# 1. Predicted probabilities
predicted_probs <- predict(aic_model, type = "response")

# 2. ROC and AUC
roc_obj <- roc(df_clean$Revenue_num, predicted_probs)
auc_val <- auc(roc_obj)

# 3. Print metrics
cat("AIC:", AIC(aic_model), "\n")
cat("AUC:", round(auc_val, 4), "\n")

# 4. Plot
plot(roc_obj, col = "#2c3e50", lwd = 3,
     main = paste("ROC Curve (Optimized Model)\nAUC =", round(auc_val, 3)))
abline(a = 0, b = 1, lty = 2, col = "red")
```
major variable

user intention data (PageValues, ExitRates)

time seryie data (Month)

user behavior data (ProductRelated_Duration, VisitorType)

"Excluded Variables":

Technical: Browser, OS, Region (Minimal impact on intent).

Redundant: SpecialDay, Weekend (Captured by Month/Seasonality).

Operational: Administrative, Informational (Less predictive than product-related engagement).


